Price of Anarchy for Mean Field Games

### 2018-07-31 PoA.tex
minor edit remark 6

\subsubsection*{\textbf{Cost Functionals}}
We assume that we are given two functions $f:[0,T] \times \mathbb{R} \times \cP(\mathbb{R}) \times \mathbb{A} \times \cP(\mathbb{A}) \ni (t,x,\mu,\alpha, \nu) \rightarrow \mathbb{R}$ and $g:\mathbb{R}\times \cP(\mathbb{R}) \ni (x,\mu) \rightarrow \mathbb{R}$ which we call running and terminal cost functions, respectively. We assume $f$ and $g$ are Lipschitz in each of their arguments. The goal of player $i$ is to minimize their expected cost as given by:
\begin{equation*}
J^i(\balpha^1,\cdots,\balpha^N)=\EE\bigg[\int_0^Tf(t,X^i_t,\mu^N_{t},\alpha^i_t,\nu^N_{t})\,dt +g(X^i_T,\mu^{N}_{T})\bigg].
\end{equation*}

@@

\subsection{\textbf{Linear Quadratic Control of McKean-Vlasov Type Involving the Law of the Control}}\label{sec:EMKV}
To solve the linear quadratic optimal control problem of McKean-Vlasov type involving the law of the control (LQEMKV), we begin with the reduced Hamiltonian, which is the same as in the LQEMFG problem:
\begin{equation*}
\begin{split}
    H(t,x,\bar{\mu},\alpha,\bar{\nu},y)=&\left[b_1(t)x+\bar{b}_1(t)\bar{\mu}+b_2(t) \alpha+\bar{b}_2(t)\bar{\nu}\right]y\\
    &+\frac{1}{2}\left[q(t)x^2+\bar{q}(t)(x-s(t)\bar{\mu})^2 +r(t)\alpha^2+\bar{r}(t)(\alpha-\bar{s}(t)\bar{\nu})^2\right].
\end{split}
\end{equation*}
Since we require $\bar{\nu}_t$ to be equal to $\mathbb{E}(\alpha_t)$ throughout the optimization, it is not sufficient to minimize the Hamiltonian with respect to the $\alpha$ input alone in order to guarantee optimality. A sufficient condition for control problems of McKean-Vlasov type involving the law of the control is derived in \cite{carmona_acciaio}. Since we consider a Hamiltonian that depends on the means of $\bar{\mu}$ and $\bar{\nu}$ instead of the full distributions, the sufficient condition reduces to the following (see section 4 in \cite{carmona_acciaio}):
$\hat{\alpha}(t,X_t,\bar{\mu},\bar{\nu},Y_t)$ should satisfy:
\begin{equation*}
    \partial_{\alpha}H(t,X_t,\mathbb{E}(X_t),\hat{\alpha}_t,\mathbb{E}(\hat{\alpha}_t),Y_t)+\tilde{\mathbb{E}} \left[\partial_{\bar{\nu}}H(t,\tilde{X}_t,\mathbb{E}(X_t),\hat{\alpha}_t,\mathbb{E}(\hat{\alpha}_t),\tilde{Y}_t) \right]=0,
\end{equation*}
where the adjoint equation is given by:
\begin{equation*}
\begin{split}
    dY_t&=-\left[\partial_x H(t,X_t,\bar{\mu}_t,\alpha_t,\bar{\nu}_t,Y_t)+\tilde{\mathbb{E}} \left[\partial_{\bar{\mu}}H(t,\tilde{X}_t,\bar{\mu}_t,\tilde{\alpha}_t,\bar{\nu}_t,\tilde{Y}_t) \right] \right]dt+Z_tdW_t \\
    Y_T&=\partial_xg(X_T, \cL(X_T))+\tilde{\mathbb{E}} \left[\partial_{\bar{\mu}}g(\tilde{X}_T,\cL(X_T))(X_T) \right],
\end{split}
\end{equation*}
and $(\boldsymbol{\tilde{X}},\boldsymbol{\tilde{Y}},\boldsymbol{\tilde{\alpha}})$ denotes an independent copy of $(\boldsymbol{X},\boldsymbol{Y},\boldsymbol{\alpha})$. In the present LQ case, the sufficient condition can be used to solve for:
\begin{equation*}
    \hat{\alpha}_t=a^{MKV}(t)Y_t+b^{MKV}(t)\mathbb{E}(Y_t),
\end{equation*}
and:
\begin{equation*}
    \mathbb{E}(\hat{\alpha}_t)=c^{MKV}(t)\mathbb{E}(Y_t),
\end{equation*}
with:
\begin{equation*}
\begin{split}
    a^{MKV}(t)&=-\frac{b_2(t)}{r(t)+\bar{r}(t)} \\
    b^{MKV}(t)&=-\frac{1}{r(t)+\bar{r}(t)}\left(\bar{b}_2(t)-\frac{\bar{r}(t)\bar{s}(t)(\bar{s}(t)-2)(b_2(t)+\bar{b}_2(t))}{r(t)+\bar{r}(t)(1-\bar{s}(t))^2} \right) \\
    c^{MKV}(t)&=-\frac{b_2(t)+\bar{b}_2(t)}{r(t)+\bar{r}(t)(1-\bar{s}(t))^2}.
\end{split}
\end{equation*}
Note that $c^{MKV}(t)=a^{MKV}(t)+b^{MKV}(t)$. So the solution of the optimal control problem of McKean-Vlasov type is given by the solution to the FBSDE system:
\begin{equation}
\begin{split}
        dX_t=&\left[b_1(t)X_t+\bar{b}_1(t) \mathbb{E}X_t+a^{MKV}(t)b_2(t)Y_t+(b^{MKV}(t)b_2(t)+c^{MKV}(t)\bar{b}_2(t))\mathbb{E}Y_t\right]dt+\sigma dW_t \\
        dY_t=&-\left[(q(t)+\bar{q}(t))X_t+s(t)\bar{q}(t)(s(t)-2)\mathbb{E}X_t+b_1(t)Y_t+\bar{b}_1(t)\mathbb{E}Y_t\right]dt +Z_t dW_t,
\end{split}
\label{eq:FBSDE_EMKV}
\end{equation}
with initial condition $X_0=\xi$, and terminal condition $Y_T=(q_T+\bar{q}_T)X_T+s_T\bar{q}_T(s_T-2)\mathbb{E}X_T$.

@@

\begin{remark}
	Consider assumption \ref{assumption} and the case when $\bar{b}_2$ tends to zero. We have $0< SC^{MKV, \bar{b}_2 \to 0} < \infty$, and therefore, $\lim_{\bar{b}_2 \to 0} PoA^{\bar{b}_2 } =1$ implies $\lim_{\bar{b}_2 \to 0} \Delta SC^{\bar{b}_2} = 0$. Hence, from equation (\ref{eq:x_bar_neq_0}) and Proposition \ref{prop:diff_SC}, if $\lim_{\bar{b}_2 \to 0} PoA^{\bar{b}_2 } =1$, then for all $t \in [0,T]$:
	$$u_t^{\bar{b}_2 \to 0} = w_t^{\bar{b}_2 \to 0}.$$
	From equation (\ref{eq:riccati_limiting_u}) and an analogous Riccati equation for $w_t^{\bar{b}_2 \to 0}$, the above equality is equivalent to:
	$$D^{u,\bar{b}_2 \to \infty} = D^w=:D \quad \text{and} \quad B^{\bar{b}_2 \to 0} D^2 + 2 A^{u}D^{u,\bar{b}_2 \to 0} - C^{u,\bar{b}_2 \to 0} = B^{\bar{b}_2 \to 0} D^2 + 2 A^{w} D^w - C^w  = 0.$$
	In other words, the following three equations together imply $\lim_{\bar{b}_2 \to 0} PoA^{\bar{b}_2 } =1$:
	$$	\frac{r + \bar{r}(1- \bar{s})^2}{r + \bar{r}(1-\bar{s})} = \frac{q_T+\bar{q}_T(1-s_T)^2}{q_T+\bar{q}_T(1-s_T)},
	\qquad
	 \bar{b}_1 \frac{q_T + \bar{q}_T(1-s_T)^2}{q + \bar{q}(1-s)^2} =  \frac{r + \bar{r}(1- \bar{s})^2}{r + \bar{r}(1-\bar{s})}\cdot \frac{q + \bar{q}(1-s)}{q + \bar{q}(1-s)^2}-1,
	$$
	and
	$$ \frac{b_2^2}{r + \bar{r}(1-\bar{s})^2} \left(q_T + \bar{q}_T(1- s_T)^2 \right)^2 - 2 (b_1 + \bar{b}_1) \left(q_T + \bar{q}_T(1- s_T)^2 \right) - \left(q + \bar{q}(1- s)^2 \right) = 0.$$\\
\label{remark_6}
\end{remark}

@@

	Since $\lim_{\bar{b}_1 \to \infty}B (D^{u})^2 + 2 A^{u,\bar{b}_1} D^{u} - C^{u}=-\infty$ and $\lim_{\bar{b}_1 \to \infty}B (D^{w})^2 + 2 A^{w,\bar{b}_1} D^{w} - C^{w}=-\infty$,   from equation (\ref{eq:riccati_ut_sol_deriv}) there exists a $\bar{b}_1^{*,lower}$ such that for $\bar{b}_1 \geq \bar{b}_1^{*,lower}$, we have:
	$$\max \left\{ \left|\frac{u^{\bar{b}_1}_0}{\bar{b}_1}-\frac{1}{B} \right|,\left|\frac{w^{\bar{b}_1}_{T/2}}{\bar{b}_1}-\frac{2}{B} \right| \right\} <\frac{1}{4B},$$
	and such that the functions $t \mapsto u_t^{\bar{b}_1}$ and $ t \mapsto w_t^{\bar{b}_1}$ are decreasing. Thus, for all $t \in [0, T/2]$ we have:
	\begin{equation*}
	    0< \frac{u_t^{\bar{b}_1}}{\bar{b}_1} \leq \frac{u_0^{\bar{b}_1}}{\bar{b}_1} \leq \frac{1}{B} + \frac{1}{4B}  \leq \frac{2}{B} - \frac{1}{4 B} \leq \frac{w_{T/2}^{\bar{b}_1}}{ \bar{b}_1} \leq \frac{w_t^{\bar{b}_1}}{\bar{b}_1}. 
	\end{equation*}
	Thus, for all $\bar{b}_1 \geq \bar{b}_1^{*,lower}$ and all $t \in [0,T/2]$,
	\begin{equation}
	  \left\vert  \frac{w^{\bar{b}_1}_t}{\bar{b}_1}-\frac{u^{\bar{b}_1}_t}{\bar{b}_1} \right\vert \geq \frac{1}{2B}.
	\label{eq:lower_bound_u_w}
	\end{equation}

>>>>

\subsubsection*{\textbf{Cost Functionals}}
We assume that we are given two functions $f:[0,T] \times \mathbb{R} \times \cP(\mathbb{R}) \times \mathbb{A} \times \cP(\mathbb{A}) \ni (t,x,\mu,\alpha, \nu) \rightarrow \mathbb{R}$ and $g:\mathbb{R}\times \cP(\mathbb{R}) \ni (x,\mu) \rightarrow \mathbb{R}$ which we call running and terminal cost functions, respectively. We assume $f$ and $g$ are Lipschitz in each of their arguments. The goal of player $i$ is to minimize its expected cost as given by:
\begin{equation*}
J^i(\balpha^1,\cdots,\balpha^N)=\EE\bigg[\int_0^Tf(t,X^i_t,\mu^N_{t},\alpha^i_t,\nu^N_{t})\,dt +g(X^i_T,\mu^{N}_{T})\bigg].
\end{equation*}

@@

\subsection{\textbf{Linear Quadratic Control of McKean-Vlasov Type Involving the Law of the Control}}\label{sec:EMKV}
To solve the linear quadratic optimal control problem of McKean-Vlasov type involving the law of the control (LQEMKV), we begin with the reduced Hamiltonian, which is the same as in the LQEMFG problem:
\begin{equation*}
\begin{split}
    H(t,x,\bar{\mu},\alpha,\bar{\nu},y)=&\left[b_1(t)x+\bar{b}_1(t)\bar{\mu}+b_2(t) \alpha+\bar{b}_2(t)\bar{\nu}\right]y\\
    &+\frac{1}{2}\left[q(t)x^2+\bar{q}(t)(x-s(t)\bar{\mu})^2 +r(t)\alpha^2+\bar{r}(t)(\alpha-\bar{s}(t)\bar{\nu})^2\right].
\end{split}
\end{equation*}
Since we require $\bar{\nu}_t$ to be equal to $\mathbb{E}(\alpha_t)$ throughout the optimization, it is not sufficient to minimize the Hamiltonian with respect to the $\alpha$ input alone in order to guarantee optimality. A sufficient condition for control problems of McKean-Vlasov type involving the law of the control is derived in \cite{carmona_acciaio}. Since we consider a Hamiltonian that depends on the means of $\bar{\mu}$ and $\bar{\nu}$ instead of the full distributions, the sufficient condition reduces to the following (see section 4 in \cite{carmona_acciaio}):
$\hat{\alpha}(t,X_t,\bar{\mu},\bar{\nu},Y_t)$ should satisfy:
\begin{equation*}
    \partial_{\alpha}H(t,X_t,\mathbb{E}(X_t),\hat{\alpha}_t,\mathbb{E}(\hat{\alpha}_t),Y_t)+\tilde{\mathbb{E}} \left[\partial_{\bar{\nu}}H(t,\tilde{X}_t,\mathbb{E}(X_t),\hat{\alpha}_t,\mathbb{E}(\hat{\alpha}_t),\tilde{Y}_t) \right]=0,
\end{equation*}
where the adjoint equation is given by:
\begin{equation*}
\begin{split}
    dY_t&=-\left[\partial_x H(t,X_t,\bar{\mu}_t,\alpha_t,\bar{\nu}_t,Y_t)+\tilde{\mathbb{E}} \left[\partial_{\bar{\mu}}H(t,\tilde{X}_t,\bar{\mu}_t,\tilde{\alpha}_t,\bar{\nu}_t,\tilde{Y}_t) \right] \right]dt+Z_tdW_t \\
    Y_T&=\partial_xg(X_T, \cL(X_T))+\tilde{\mathbb{E}} \left[\partial_{\bar{\mu}}g(\tilde{X}_T,\cL(X_T))(X_T) \right],
\end{split}
\end{equation*}
and $(\boldsymbol{\tilde{X}},\boldsymbol{\tilde{Y}},\boldsymbol{\tilde{\alpha}})$ denotes an independent copy of $(\boldsymbol{X},\boldsymbol{Y},\boldsymbol{\alpha})$. In the present LQ case, the sufficient condition can be used to solve for:
\begin{equation*}
    \hat{\alpha}_t=a^{MKV}(t)Y_t+b^{MKV}(t)\mathbb{E}(Y_t),
\end{equation*}
and:
\begin{equation*}
    \mathbb{E}(\hat{\alpha}_t)=c^{MKV}(t)\mathbb{E}(Y_t),
\end{equation*}
with:
\begin{equation*}
\begin{split}
    a^{MKV}(t)&=-\frac{b_2(t)}{r(t)+\bar{r}(t)} \\
    b^{MKV}(t)&=-\frac{1}{r(t)+\bar{r}(t)}\left(\bar{b}_2(t)-\frac{\bar{r}(t)\bar{s}(t)(\bar{s}(t)-2)(b_2(t)+\bar{b}_2(t))}{r(t)+\bar{r}(t)(1-\bar{s}(t))^2} \right) \\
    c^{MKV}(t)&=-\frac{b_2(t)+\bar{b}_2(t)}{r(t)+\bar{r}(t)(1-\bar{s}(t))^2}.
\end{split}
\end{equation*}
So the solution of the optimal control problem of McKean-Vlasov type is given by the solution to the FBSDE system:
\begin{equation}
\begin{split}
        dX_t=&\left[b_1(t)X_t+\bar{b}_1(t) \mathbb{E}X_t+a^{MKV}(t)b_2(t)Y_t+(b^{MKV}(t)b_2(t)+c^{MKV}(t)\bar{b}_2(t))\mathbb{E}Y_t\right]dt+\sigma dW_t \\
        dY_t=&-\left[(q(t)+\bar{q}(t))X_t+s(t)\bar{q}(t)(s(t)-2)\mathbb{E}X_t+b_1(t)Y_t+\bar{b}_1(t)\mathbb{E}Y_t\right]dt +Z_t dW_t,
\end{split}
\label{eq:FBSDE_EMKV}
\end{equation}
with initial condition $X_0=\xi$, and terminal condition $Y_T=(q_T+\bar{q}_T)X_T+s_T\bar{q}_T(s_T-2)\mathbb{E}X_T$.

@@

\begin{remark}
	Consider assumption \ref{assumption} and the case when $\bar{b}_2$ tends to zero. We have $0< SC^{MKV, \bar{b}_2 \to 0} < \infty$, and therefore, $\lim_{\bar{b}_2 \to 0} PoA^{\bar{b}_2 } =1$ implies $\lim_{\bar{b}_2 \to 0} \Delta SC^{\bar{b}_2} = 0$. Hence, from equation (\ref{eq:x_bar_neq_0}) and Proposition \ref{prop:diff_SC}, if $\lim_{\bar{b}_2 \to 0} PoA^{\bar{b}_2 } =1$, then for all $t \in [0,T]$:
	$$u_t^{\bar{b}_2 \to 0} = w_t^{\bar{b}_2 \to 0}.$$
	From equation (\ref{eq:riccati_limiting_u}) and an analogous Riccati equation for $w_t^{\bar{b}_2 \to 0}$, the above equality is equivalent to:
	$$D^{u,\bar{b}_2 \to \infty} = D^w=:D \quad \text{and} \quad B^{\bar{b}_2 \to 0} D^2 + 2 A^{u}D - C^{u,\bar{b}_2 \to 0} = B^{\bar{b}_2 \to 0} D^2 + 2 A^{w} D - C^w  = 0.$$
	In other words, the following three equations together imply $\lim_{\bar{b}_2 \to 0} PoA^{\bar{b}_2 } =1$:
	$$	\frac{r + \bar{r}(1- \bar{s})^2}{r + \bar{r}(1-\bar{s})} = \frac{q_T+\bar{q}_T(1-s_T)^2}{q_T+\bar{q}_T(1-s_T)},
	\qquad
	 \bar{b}_1 \frac{q_T + \bar{q}_T(1-s_T)^2}{q + \bar{q}(1-s)^2} =  \frac{r + \bar{r}(1- \bar{s})^2}{r + \bar{r}(1-\bar{s})}\cdot \frac{q + \bar{q}(1-s)}{q + \bar{q}(1-s)^2}-1,
	$$
	and
	$$ \frac{b_2^2}{r + \bar{r}(1-\bar{s})^2} \left(q_T + \bar{q}_T(1- s_T)^2 \right)^2 - 2 (b_1 + \bar{b}_1) \left(q_T + \bar{q}_T(1- s_T)^2 \right) - \left(q + \bar{q}(1- s)^2 \right) = 0.$$\\
\label{remark_6}
\end{remark}

@@

	Since $\lim_{\bar{b}_1 \to \infty}B (D^{u})^2 + 2 A^{u,\bar{b}_1} D^{u} - C^{u}=-\infty$ and $\lim_{\bar{b}_1 \to \infty}B (D^{w})^2 + 2 A^{w,\bar{b}_1} D^{w} - C^{w}=-\infty$,   from equation (\ref{eq:riccati_ut_sol_deriv}) there exists a $\bar{b}_1^{*,lower}$ such that for $\bar{b}_1 \geq \bar{b}_1^{*,lower}$, we have:
	$$\max \left\{ \left|\frac{u^{\bar{b}_1}_0}{\bar{b}_1}-\frac{1}{B} \right|,\left|\frac{w^{\bar{b}_1}_{T/2}}{\bar{b}_1}-\frac{2}{B} \right| \right\} \leq \frac{1}{4B},$$
	and such that the functions $t \mapsto u_t^{\bar{b}_1}$ and $ t \mapsto w_t^{\bar{b}_1}$ are decreasing. Thus, for all $t \in [0, T/2]$ we have:
	\begin{equation*}
	    0< \frac{u_t^{\bar{b}_1}}{\bar{b}_1} \leq \frac{u_0^{\bar{b}_1}}{\bar{b}_1} \leq \frac{1}{B} + \frac{1}{4B}  \leq \frac{2}{B} - \frac{1}{4 B} \leq \frac{w_{T/2}^{\bar{b}_1}}{ \bar{b}_1} \leq \frac{w_t^{\bar{b}_1}}{\bar{b}_1}. 
	\end{equation*}
	Thus, for all $\bar{b}_1 \geq \bar{b}_1^{*,lower}$ and all $t \in [0,T/2]$,
	\begin{equation}
	  \left\vert  \frac{w^{\bar{b}_1}_t}{\bar{b}_1}-\frac{u^{\bar{b}_1}_t}{\bar{b}_1} \right\vert \geq \frac{1}{2B}.
	\label{eq:lower_bound_u_w}
	\end{equation}
